---
title: "data_testing"
author: "Ngoc Duong - nqd2000"
date: "2/8/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE}
library(tidyverse)
library(MASS)
library(matrixcalc)
library(ggplot2)
```

Write function to simulate dataset with 3 kinds of given predictors + null predictors
```{r}
sim_beta_strong = function(n_strong, coef_strong){
  rep(coef_strong, n_strong) + runif(n_strong, min = 0, max = coef_strong)
}


sim_data = function(n_parameter, n_sample, prop_strong = 0.1, prop_wbc = 0.2, prop_wai = 0.2, c = 1, cor = 0.30, coef_strong = 10) {
  
  # Numbers of four signals
  n_strong = as.integer(n_parameter * prop_strong) # strong
  n_wbc = as.integer(n_parameter * prop_wbc) # weak but correlated
  n_wai = as.integer(n_parameter * prop_wai) # weak and independent
  n_null = n_parameter - n_strong - n_wbc - n_wai # null
    
  if (n_null < 0) {
    return("Given parameters' proportions are not valid.")
  }
  
  bound = c * sqrt(log(n_parameter) / n_sample) # threshold of weak/strong, the default is 0.14
  if (coef_strong < bound) {
    coef_strong = coef_strong + 2 * bound
  }
  
  cor_matrix = diag(n_parameter)
  # wbc covariates are correlated to the first strong covariate
  cor_matrix[1, (n_strong + n_wai + 1):(n_strong + n_wai + n_wbc)] = cor
  cor_matrix[(n_strong + n_wai + 1):(n_strong + n_wai + n_wbc), 1] = cor
  
  if (!is.positive.definite(cor_matrix)) {
    return("The correlation matrix is not valid.")
  }
  
  # simulate the data from multivariate normal
  X = mvrnorm(n = n_sample, mu = rep(0, n_parameter), Sigma = cor_matrix) # var = 1, correlation = covariance
  
  beta = c(
    sim_beta_strong(n_strong, coef_strong),
    runif(min = bound/2, max = bound, n = n_wai), 
    runif(min = bound/2, max = bound, n = n_wbc),
    rep(0, n_null) 
  )
  
  Y = 1 + X %*% beta + rnorm(n_sample)
  data = as_tibble(data.frame(cbind(X, Y)))
  
  # Name the columns
  cols = c(
    str_c("strong", 1:n_strong, sep = "_"),
    str_c("wai", 1:n_wai, sep = "_"),
    str_c("wbc", 1:n_wbc, sep = "_"),
    str_c("null", 1:n_null, sep = "_"),
    "Y"
   )
   colnames(data) = cols
   data = data %>% 
     dplyr::select(Y, everything())
}

```

Function implementing forward selection method using AIC as criterion
```{r}
forward.aic.lm = function(df) {
  null.lm = lm(Y ~ 1, data = df)
  full.lm = lm(Y ~ ., data = df)
  aic.lm = step(object = null.lm,
             scope = list(lower = null.lm, 
                          upper = full.lm), 
             direction = "forward",
             trace = FALSE, 
             k = 2)
  aic.lm
}
```


Write function to calculate:
1) how many strong predictors are selected by model
2) how many percent of strong predictors are selected by model
3) how many wbi and wac predictors are missed my the model 
4) how many percent of wbi and wac predictors are missed my the model 

```{r}
predictors.analysis = function(model_coeff, n_parameter,
                               prop_strong = 0.1, prop_wbc = 0.2, prop_wai = 0.2) {
  
  n_strong = as.integer(n_parameter * prop_strong) # strong
  n_wbc = as.integer(n_parameter * prop_wbc) # weak but correlated
  n_wai = as.integer(n_parameter * prop_wai) # weak and independent
  n_null = n_parameter - n_strong - n_wbc - n_wai # number of null
  model_coeff = names(model_coeff)
  return(tibble(
    n_parameter_total = n_parameter,
    n_parameter_selected = length(model_coeff)-1,
    n_strong_selected = length(which(str_detect(model_coeff, "strong"))),
    n_wai_selected = length(which(str_detect(model_coeff, "wai"))),
    n_wbc_selected = length(which(str_detect(model_coeff, "wbc"))),
    n_null_selected =  n_parameter_selected - n_strong_selected - n_wai_selected - n_wbc_selected,
    prop_strong =  round(n_strong_selected/n_strong,2),
    prop_wai = round(n_wai_selected/n_wai,2),
    prop_wbc = round(n_wbc_selected/n_wbc,2),
    prop_null = round(n_null_selected/n_null,2)
  ))
}

```




Scenario 1: while changing number of parameters, simulate the data 100 times for each parameter
```{r}
#key parameters
forward_summary = NULL
n_parameter = c(10,20,30,40,50)
n_sim = 30
n_sample = 1000

for(i in n_parameter){
    for (j in 1:n_sim) {
      df = sim_data(n_parameter = i, n_sample = n_sample)
      
    # create the forward model
      forward_lm = forward.aic.lm(df)
      
    # calculate the percentages of coefficients missed
      forward_data = predictors.analysis(model_coeff = forward_lm$coefficients, 
                                         n_parameter = i)
      # Gather created models for diagnostics
      forward_summary = rbind(forward_summary, forward_data)
    }
}
forward_summary_final = forward_summary %>% 
  group_by(n_parameter_total) %>% 
  mutate(strong_prfm = mean(prop_strong),
         wai_prfm = mean(prop_wai),
         wbc_prfm = mean(prop_wbc), 
         null_prfm = mean(prop_null))
```

Visualization with all simulations:
1) How many percent of variables are selected by method, on average?

```{r 1}
forward_summary_final %>% ggplot(aes(x = n_parameter_total))+
 geom_point(aes(y=strong_prfm*100, color="prop_strong")) +
 geom_point(aes(y=wai_prfm*100, color="prop_wai")) + 
 geom_point(aes(y=wbc_prfm*100, color="prop_wbc")) + 
 geom_point(aes(y=null_prfm*100, color="prop_null")) + 
  geom_line(aes(y=strong_prfm*100, color="prop_strong")) +
  geom_line(aes(y=wai_prfm*100, color="prop_wai")) + 
  geom_line(aes(y=wbc_prfm*100, color="prop_wbc")) + 
  geom_line(aes(y=null_prfm*100, color="prop_null")) + 
 labs(title = "Percent of each type of signal included in the forward selection model",
       y = "Percent of signals included in the model (%)",
       x = "Number of (pre-set) total parameters") +  
  theme_bw() + 
  theme(plot.title = 
          element_text(hjust = 0.5, 
                       size=12, 
                       face='bold'), 
          legend.position = "bottom") +
  scale_x_discrete(limits=c("prop_strong", "prop_wai", "prop_wbc", "prop_null")) + 
  scale_color_discrete(name = "Signal Type", labels = c("Strong", "WAI", "WBC", "Null")) 

#represent the same type of info using different visualization (but does not require taking mean)
forward_summary_final %>% ggplot(aes()) +
geom_density(aes(x = prop_strong, color = "Strong", alpha = 0.20)) +
geom_density(aes(x = prop_wbc, color = "WBC", alpha = 0.20)) + 
geom_density(aes(x=prop_wai, color = "WAI", alpha=0.20)) +
geom_density(aes(x=prop_null, color = "Null", alpha=0.20)) +
  facet_grid(~n_parameter_total) +
    theme(legend.title=element_blank())
```


2) Number of strong predictors changed by more than 10%, expressed in terms of how many weak predictors are missing
n_parameter = c(10,20,30,40,50)
n_sim = 30
n_sample = 1000

beta.analysis = function(betas){
  #calculate the number of wbc and wai missing
  
  #Calculate MSE of the strong predictors 
  
}

for(i in n_parameter){
    for (j in 1:n_sim) {
      df = sim_data(n_parameter = i, n_sample = n_sample)
    create the forward model
    forward_lm = forward.aic.lm(df)
    }     


Function to assess variance and bias of strong predictors' coefficients with respect to the number of missing weak variables


